Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-08-27 GWLDCII.tex
started section 4

In Section~\ref{sec GN algorithm} we state an algorithm for constructing the Grassmann necklace of the positroid associated to a Wilson loop diagram (Algorithm~\ref{alg:put GN on WLD}) and prove its correctness (Theorem~\ref{res:alg gives GN}). While it was already possible to construct the Grassmann necklace of a Wilson loop diagram via existing bijections between some of the various combinatorial objects that index the positroid cells, this process was convoluted and involved multiple steps; our algorithm allows us to track directly how each propagator in the Wilson loop diagram contributes to each term in the Grassmann necklace. Many of the technical lemmas required to prove Theorem~\ref{res:alg gives GN} continue to be important in subsequent sections: in particular we have Lemma~\ref{lem sian}, which identifies certain configurations of propagators that must appear in any admissible Wilson loop diagram, and Lemma~\ref{vertex cyclic int lem}, which shows that the pattern of values that a propagator contributes to each term of the Grassmann necklace is both simple and predictable.

@@

In this section we will give an algorithm (Algorithm~\ref{alg:put GN on WLD}) to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid. This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given WLD. It also allows us to examine how the positroid changes as we make minor changes to the Wilson loop diagram; indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections.

@@

\sanote{START HERE}

We can now unambiguously refer to Algorithm~\ref{alg:put GN on WLD} as the Grassmann necklace algorithm. We close this section with a simple observation about the Grassmann necklaces of positroids associated to Wilson loop diagrams.

@@

\begin{cor}\label{no coloops}\sanote{cite more prominently in frontmatter}
Grassmann necklaces coming from admissible Wilson loop diagrams have no coloops. A vertex $j$ is a loop if and only if $j$ supports no propagators. 
\end{cor}
\begin{proof}
For any $i \in [n]$, $i-1$ is maximal with respect to the $<_{i}$ order and so by Lemma~\ref{lem no fourth vertex} there can be no propagator $p$ with $I_{i}(p) = i-1$. Thus $i-1 \not\in I_{i}$, so $i-1$ is not a coloop. Since this holds for any $i \in [n]$, the Grassmann necklace admis no coloops.

@@

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$ (Theorem \ref{thm dim}).  The positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{non-orientable}, Agarwala and Marcott show that the dimension of the Deodhar component is 
\[|\cP| + \text{ the dimension of associated positroid cell}.\]
\note{What if instead of displaying this we wrote ``is the sum of $\|\cP\|$ and the dimension fo the associated positroid cell.''}
In other words, the Deodhar component associated to a Wilson loop diagram is $4|\cP|$, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \hlfix{***cite this***}{highlighted for anyone skimming through} gives a different proof of the $3|\cP|$-dimensionality of $M(W)$ which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

@@

@@

\begin{proof}
  By Algorithm~\ref{alg:put GN on WLD} the Grassmann necklace of $W$ is obtained from the Grassmann necklace of $V$ by duplicating the $i$th element of the Grassmann necklace of $V$ (shifting indices by $1$ for all indices greater than $i$), and incrementing all indices greater than $i$ by $1$ in each Grassmann necklace element\note{This is saying to increment twice in a way which I think is confusing, but presumably the parenthetical remark was put there to try to clarify something so I don't want to just remove it}.  Formally, if $I_1^{V}, \ldots, I_{n-1}^{V}$ and $I_1^{W}, \ldots, I_n^{W}$ are the Grassmann necklaces of $V$ and $W$ respectively then
  \[
  I_j^{W} =
  \begin{cases}
    \{\ell \in I_j^{V} : \ell < i\} \cup \{\ell+1 \in I_j^{V} : \ell \geq i\} & \text{if $j\leq i$} \\
    \{\ell \in I_{j-1}^{V} : \ell < i\} \cup \{\ell+1 \in I_{j-1}^{V} : \ell \geq i\} & \text{if $j > i$.}
  \end{cases}
  \]

By Lemma~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column since $I_1^*$\note{I don't find the star very helpful here; different area's conventions maybe} is the same for $V$ and $W$ except for the incrementation of the indices $\geq i$ in the transition from the necklace for $V$ to the necklace for $W$. 
\end{proof}

>>>>

In Section~\ref{sec GN algorithm} we state an algorithm for constructing the Grassmann necklace of the positroid associated to a Wilson loop diagram (Algorithm~\ref{alg:put GN on WLD}) and prove its correctness (Theorem~\ref{res:alg gives GN}). While it was already possible to construct the Grassmann necklace of a Wilson loop diagram via existing bijections between some of the various combinatorial objects that index the positroid cells, this process was convoluted and involved multiple steps; our algorithm allows us to track directly how each propagator in the Wilson loop diagram contributes to each term in the Grassmann necklace. {\color{violet} Furthermore, the above algorithm allows us to determine that the positroids associated to Wilson loop better. In particular, Corrollary \ref{no coloops} shows that these positroids have no coloops, and gives an exact characterization of their loops.}Many of the technical lemmas required to prove Theorem~\ref{res:alg gives GN} continue to be important in subsequent sections: in particular we have Lemma~\ref{lem sian}, which identifies certain configurations of propagators that must appear in any admissible Wilson loop diagram, and Lemma~\ref{vertex cyclic int lem}, which shows that the pattern of values that a propagator contributes to each term of the Grassmann necklace is both simple and predictable.

@@

In this section we will give an algorithm (Algorithm~\ref{alg:put GN on WLD}) to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid. This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given Wilson loop diagram. {\color{violet} Furthemore, we use it to characterise the positroids associated to Wilson loop diagrams according to loops and coloops (Corrollary \ref{no coloops}).} It also allows us to examine how the positroid changes as we make minor changes to the Wilson loop diagram; indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections.

@@

We have now shown that Algorithm~\ref{alg:put GN on WLD} assoiates the correct Grassmann necklace to a Wilson Loop diagam Thereofre, we refer to Algorithm~\ref{alg:put GN on WLD} as the Grassmann necklace algorithm. We close this section with a simple observation about the Grassmann necklaces of positroids associated to Wilson loop diagrams.

@@

\begin{cor}\label{no coloops}
Grassmann necklaces coming from admissible Wilson loop diagrams have no coloops. A vertex $j$ is a loop if and only if $j$ supports no propagators. 
\end{cor}
\begin{proof}
For any $i \in [n]$, $i-1$ is maximal with respect to the $<_{i}$ order and so by Lemma~\ref{lem no fourth vertex} there can be no propagator $p$ with $I_{i}(p) = i-1$. Thus $i-1 \not\in I_{i}$, so $i-1$ is not a coloop. Since this holds for any $i \in [n]$, the Grassmann necklace admits no coloops.

@@

In continuing to characterise the poistroid cells associated to Wilson loop diagrams, our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$ (Theorem \ref{thm dim}).  The positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite{non-orientable}, Agarwala and Marcott show that the dimension of the Deodhar component is the sum of $\|\cP\|$ and the dimension fo the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
\note{What if instead of displaying this we wrote ``is the sum of $\|\cP\|$ and the dimension fo the associated positroid cell.''}\sanote{implemented, but don't like how it no longer stands out. Thoughts?}
In other words, the Deodhar component associated to a Wilson loop diagram is $4|\cP|$, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \cite{WLDdim} gives a different proof of the $3|\cP|$-dimensionality of $M(W)$ which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

@@

\sanote{musing: would it be worth moving the LE diagram exposition here? Do we use it ealier? I like having exposition in front generally, but it feels a mile away.}

@@

\begin{proof}
  By Algorithm~\ref{alg:put GN on WLD} the Grassmann necklace of $W$ is obtained from the Grassmann necklace of $V$ by duplicating the $i$th element of the Grassmann necklace of $V$ (shifting indices by $1$ for all indices greater than $i$), and incrementing all indices greater than $i$ by $1$ in each Grassmann necklace element\note{This is saying to increment twice in a way which I think is confusing, but presumably the parenthetical remark was put there to try to clarify something so I don't want to just remove it}.  Formally, \st{if $I_1^{V}, \ldots, I_{n-1}^{V}$ and $I_1^{W}, \ldots, I_n^{W}$ are the Grassmann necklaces of $V$ and $W$ respectively then}\sanote{in favor of removing parenthetical, we say same thing twice without it}
  \[
  I_j^{W} =
  \begin{cases}
    \{\ell \in I_j^{V} : \ell < i\} \cup \{\ell+1 \in I_j^{V} : \ell \geq i\} & \text{if $j\leq i$} \\
    \{\ell \in I_{j-1}^{V} : \ell < i\} \cup \{\ell+1 \in I_{j-1}^{V} : \ell \geq i\} & \text{if $j > i$.}
  \end{cases}
  \]

By Corrollary~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column since $I_1^*$\note{I don't find the star very helpful here; different area's conventions maybe} is the same for $V$ and $W$ except for the incrementation of the indices $\geq i$ in the transition from the necklace for $V$ to the necklace for $W$. 
\end{proof}
